---
output: html_document
runtime: shiny
---

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup-workspace, include=FALSE}

# TO DO
## Understand github
## link to spider dat (will need real 2022 data to do so)
##Update the plots for the new veg student activity 1.4

# load all required packages
library(DT)
library(lubridate)
library(ragg)
library(readxl)
library(rstatix)
library(shiny)
library(shinyjs) 
library(writexl)
library(tidyverse) # always last to avoid conflicts

# triadic feature colours
triL_col <- "#7D31ED" # purple
triM_col <- "#ED7D31" # orange
triR_col <- "#13d863" # green -- adjusted darker 

# negated inclusive operator
"%!in%" <- Negate("%in%")

# prevent r errors etc. appearing to user
knitr::opts_chunk$set(echo = FALSE)

# set filepath root: is doc being loaded as a Shiny app or run interactively?
if (interactive() == TRUE) {
  picky_path = "biol2015/"
} else {
  picky_path = ""
}
```

```{r import-data, include=FALSE}

# import veg masterdata -- multiple sheets
# get all sheets
sheet_list <- excel_sheets(paste0(picky_path,
                                  "data/raw/M2-A1_2017-2023.xlsx"))

all_sheets <- map(.x = sheet_list,
                        .f = function(.x) {
                          read_excel(paste0(picky_path,
                                            "data/raw/M2-A1_2017-2023.xlsx"),
                                     sheet = .x,
                                     na = c("", "N/A"))
                        })

all_sheets_named <- set_names(all_sheets,
                              paste0("raw_", sheet_list))

# add imported sheets to global environment, 
# using their list names
list2env(all_sheets_named,
         envir = .GlobalEnv)

# repeat for spiders -- terrible form here, but time is of the essence
sheet_list <- excel_sheets(paste0(picky_path,
                                  "data/raw/M2-A2_2019-2023.xlsx"))

all_sheets <- map(.x = sheet_list,
                  .f = function(.x) {
                    read_excel(paste0(picky_path,
                                      "data/raw/M2-A2_2019-2023.xlsx"),
                               sheet = .x,
                               na = c("", "N/A"))
                  })

all_sheets_named <- set_names(all_sheets,
                              paste0("raw_", sheet_list))

# add imported sheets to global environment, 
# using their list names
list2env(all_sheets_named,
         envir = .GlobalEnv)

# tidy datasets as required -- most do not need tidying

# add group names from burrow_diam to burrow_count using quad_year identifier
quad_name_ref <- raw_burrow_diameter %>% 
  select(quad_year, group_name, area) %>% 
  distinct(quad_year,
           .keep_all = TRUE)

# and sum burrow counts, across any multiple surveys within each group name
burrow_count <- raw_burrow_count %>% 
  select(-group_name) %>% # trash col in this dataset
  left_join(., quad_name_ref,
            by = "quad_year") %>% 
  mutate(burrow_count = 100 / area * burrow_count)

# remove hanger-on NAs from bottom of veg_metadata
veg_metadata <- raw_veg_metadata %>% 
  filter(!is.na(variable))

spider_metadata <- raw_spider_metadata

# full metadata reference for plotting code
## MOVED TO BELOW with processed_metadata

sac <- raw_species_area_curve %>% 
  filter(quadrat_area == 256)

small_quadrats <- raw_small_quadrats %>% 
  group_by(year, group_name, 
           lat, lon, alt) %>% 
  summarise(across(c(foliage_projective_cover,
                     living_ground_cover,
                     litter_cover),
                   mean, na.rm = TRUE)) %>% 
  ungroup()
  
us_estimates <- raw_understorey_estimates %>% 
  group_by(year, group_name, 
           lat, lon, alt) %>% 
  summarise(across(c(small_trees,
                     `small_shrub`,
                     `big_shrub`),
                   mean, na.rm = TRUE)) %>% 
  ungroup() 

tall_trees <- raw_tallest_trees %>% 
  group_by(year, group_name, 
           lat, lon, alt) %>% 
  summarise(across(c(circum, dbh,
                     basal_area,
                     height),
                   mean, na.rm = TRUE)) %>% 
  ungroup() 

# doesn't matter because burrow diameter is no longer offered for plotting
# but ANYWAY...
burrow_diam <- raw_burrow_diameter %>% 
  group_by(group_name) %>% 
  mutate(burrow_diam = mean(burrow_diam, na.rm = TRUE),
         lat = mean(lat_red, na.rm = TRUE),
         lon = mean(lon_red, na.rm = TRUE)) %>%
  ungroup() %>% 
  distinct(quad_year, .keep_all = TRUE) %>% 
  group_by(group_name) %>% 
  mutate(area = sum(area)) %>% 
  ungroup() %>% 
  distinct(group_name, .keep_all = TRUE) %>% 
  select(-c(year, date, quadrat, quad_year,
            lat_red, lon_red, coords_site))

# we do need this -- because historic quadrats under the same group name
# are not necessarily together -- may be far apart along the walk
burrow <- left_join(burrow_count, tall_trees,
                    by = "group_name",
            suffix = c("", ".remove")) %>%
  left_join(., us_estimates,
            by = "group_name",
            suffix = c("", ".remove")) %>%
  left_join(., small_quadrats,
            by = "group_name",
            suffix = c("", ".remove")) %>%
  left_join(., sac,
            by = "group_name",
            suffix = c("", ".remove")) %>%
  select(-c(contains(".remove"), date, 
            quadrat, area, quadrat_area))

# # save tidied veg datasets
# write_xlsx(list(tallest_trees = tall_trees,
#                 understorey_estimates = us_estimates,
#                 small_quadrats = small_quadrats,
#                 species_area_curve = sac),
#            "biol2015/data/output/2023_quadrat_level_veg.xlsx",
#            format_headers = FALSE)

# save tidied spider data
# write_xlsx(list(burrows = burrow),
#            "biol2015/data/output/2023_quadrat_level_spiders.xlsx",
#            format_headers = FALSE)

# worked offline to combine, linking 2022 sites only
# load that neater file now
processed_data_veg <- read_excel("data/output/processed_2023_M2-A1-A2_data.xlsx",
                             sheet = "veg_data") %>% 
  mutate(year = as_factor(year))

processed_data_spiders <- read_excel("data/output/processed_2023_M2-A1-A2_data.xlsx",
                             sheet = "spider_data") %>% 
  mutate(year = as_factor(year))

processed_metadata <- read_excel("data/output/processed_2023_M2-A1-A2_data.xlsx",
                             sheet = "comb_metadata_key") %>% 
  filter(dataset != "no_show")

metadata <- bind_rows(veg_metadata, spider_metadata) %>% 
  left_join(processed_metadata, .,
            by = "variable") %>% 
  distinct(variable, .keep_all = TRUE)

# create option lists for selectInput()s
choices_prep <- function (data) {
  cont_pred <- c(data %>% 
                   filter(cate_cont_type == "cont" &
                            pred_resp_type %in% c("pred", "both")) %>% 
                   pull(var = variable,
                        name = illegal_variable))
  
  cate_pred <- c(data %>% 
                   filter(cate_cont_type == "cate" &
                            pred_resp_type %in% c("pred", "both")) %>% 
                   pull(var = variable,
                        name = illegal_variable))
  
  cont_resp <- c(data %>% 
                   filter(cate_cont_type == "cont" &
                            pred_resp_type %in% c("resp", "both")) %>% 
                   pull(var = variable,
                        name = illegal_variable))
  
  return(list(cont_pred, cate_pred, cont_resp))
}

veg_choices <- choices_prep(veg_metadata)
spider_choices <- choices_prep(spider_metadata)

cont_preds <- list(" " = list("Nothing selected" = "not_sel"),
                   "Veg variables: " = veg_choices[[1]],
                   "Spider variables: " = spider_choices[[1]])

cate_preds <- list(" " = list("Nothing selected" = "not_sel"),
                   "Veg variables: " = veg_choices[[2]],
                   "Spider variables: " = spider_choices[[2]])

cont_resps <- list(" " = list("Nothing selected" = "not_sel"),
                   "Veg variables: " = veg_choices[[3]],
                   "Spider variables: " = spider_choices[[3]])

```

# Module 2 --- soil, vegetation and spiders
### BIOL2015 Field Ecology
###### v1.7 | 28/06/2023 | jwh  

## {.tabset}

### 1. Design a figure

Select a predictor (x axis) and response (y axis) variable to plot a figure. You can either use a continuous or categorical predictor. 

``` {r design-plot}

# select: choose a predictor (x) and response (y) var for plot
inputPanel(
  selectInput(inputId = "cont_pred_chce",
              label = "Continuous predictors...",
              choices = cont_preds),
  selectInput(inputId = "cate_pred_chce",
              label = "Categorical predictors...",
              choices = cate_preds),
  selectInput(inputId = "cont_resp_chce",
              label = "Continuous responses...",
              choices = cont_resps),
)

# make a reactive values object
rv <- reactiveValues()

# initialise plot variables as not selected
rv$x_term <- "not_sel"
rv$y_term <- "not_sel"

# and first time flag for the disclaimer
rv$first_time <- TRUE

```

```{r build-relevant-dat}
  
# create a tibble of relevant data for plotting/modelling use
relevant_dat <- reactive({
  if (rv$x_term == "burrow_count" | rv$y_term == "burrow_count") {
    processed_data_spiders
  } else {
    processed_data_veg
  }
  
  # processed_data
})

```

```{r form-logic}

# useshinyjs() needs to be in each chunk that uses shinyjs functions
useShinyjs(rmd = TRUE)

# observe({}) defines some actions that should be taken on some trigger;
# then piped to bindEvent() which defines the trigger to watch
observe({
  if (input$cont_pred_chce != "not_sel") {
    disable(id = "cate_pred_chce")
  } else {
    enable(id = "cate_pred_chce")
  }
  rv$x_term <- c(input$cont_pred_chce)
  rv$cont_or_cate <- "cont"
}) %>% 
  bindEvent(input$cont_pred_chce)

observe({
  if (input$cate_pred_chce != "not_sel") {
    disable(id = "cont_pred_chce")
  } else {
    enable(id = "cont_pred_chce")
  }
  rv$x_term <- c(input$cate_pred_chce)
  rv$cont_or_cate <- "cate"
}) %>% 
  bindEvent(input$cate_pred_chce)

reactive({rv$y_term <- c(input$cont_resp_chce)})

```

```{r make-plot}

observe({

  if (rv$x_term != "not_sel" & rv$y_term != "not_sel") {
    x_term <- sym(rv$x_term)
    y_term <- sym(rv$y_term)
    
    # determine axes labels from metadata reference
    x_lab <- metadata %>% 
      filter(variable == rv$x_term) %>% 
      mutate(label = paste0(illegal_variable,
                            " (", units, ")")) %>% 
      pull(label)
    
    y_lab <- metadata %>% 
      filter(variable == rv$y_term) %>% 
      mutate(label = paste0(illegal_variable,
                            " (", units, ")")) %>% 
      pull(label)
    
    # data for background plotting 
    raw_plot_dat <- relevant_dat() %>% 
      select(!!x_term, !!y_term) %>%
      filter(is.na(!!x_term) == FALSE,
             is.na(!!y_term) == FALSE)
    
    # continuous predictor; continuous response
    if (rv$cont_or_cate == "cont") {
      
      # make plot of selected variables
      selected_plot <- ggplot() +
        geom_point(data = raw_plot_dat,
                   aes(x = !!x_term,
                       y = !!y_term),
                   colour = "grey40") +
        geom_smooth(data = raw_plot_dat,
                   aes(x = !!x_term,
                       y = !!y_term),
                   colour = "black",
                   fill = "grey60",
                   method = "lm") +
        scale_x_continuous(name = x_lab) +
        scale_y_continuous(name = y_lab) +
        theme_classic()
    }
    
    # categorical predictor; continuous response
    if (rv$cont_or_cate == "cate") {
      
      # data for foreground plotting
      summary_plot_dat <- raw_plot_dat %>% 
        group_by(!!x_term) %>% 
        summarise(plot_mean = mean(!!y_term),
                  plot_se = sd(!!y_term)/sqrt(n())) %>% 
        ungroup()
      
      # make plot of selected variables
      selected_plot <- ggplot() +
        geom_point(data = raw_plot_dat,
                   aes(x = !!x_term,
                       y = !!y_term),
                   colour = "grey80",
                   position = position_jitter(width = 0.2,
                                              height = 0,
                                              seed = 2015)) +
        geom_errorbar(data = summary_plot_dat,
                      aes(x = !!x_term,
                          ymin = plot_mean - plot_se,
                          ymax = plot_mean + plot_se),
                      colour = "black",
                      width = 0,
                      size = 0.75) +
        geom_point(data = summary_plot_dat,
                   aes(x = !!x_term,
                       y = plot_mean),
                   colour = "black",
                   size = 2.5) +
        scale_x_discrete(name = x_lab) +
        scale_y_continuous(name = y_lab) +
        theme_classic()
      
      rv$summary_plot_dat <- summary_plot_dat
    }
   
    # prepare data for display
    output$table_summary_plot_dat <- renderDT({
      datatable(summary_plot_dat,
                rownames = FALSE,
                options = list(dom = "t"))
    })
    
    output$table_raw_plot_dat <- renderDT({
      datatable(raw_plot_dat,
                rownames = FALSE,
                options = list(dom = "ltp"))
    })
    
    # prepare plot for display
    output$plot_selected_plot <- renderPlot({
      selected_plot
    })
    
    # prepare same plot for printing into ragg save device
    rv$plot_selected_plot <- selected_plot
    
    # ditto for datasets and their save buttons
    # (summary done in cate-pred if, above)
    rv$raw_plot_dat <- raw_plot_dat
    
  }
}) %>% bindEvent(input$cont_pred_chce,
                 input$cate_pred_chce,
                 input$cont_resp_chce)

```

```{r show-plot}

plotOutput("plot_selected_plot")

```

```{r save-plot}

# if using downloadButton outside of the inputPanel, outside brackets
# are req'd around the function to make it render in Shiny Rmd

inputPanel(textInput("fig_name",
                     label = "Enter your first and last names for a unique filename",
                     placeholder = "e.g. MatthewLuskin"),
           # textInput("fig_width",
           #           "Enter a figure width (cm)",
           #           value = 8,
           #           placeholder = "I suggest 8"),
           # textInput("fig_height",
           #           "Enter a figure height (cm)",
           #           value = 8,
           #           placeholder = "I suggest 8"),
           downloadButton("save_plot",
                          "Save your created figure..."))

reactive({
  # fig_width <- c(input$fig_width)
  # fig_height <- c(input$fig_height)
  
  output$save_plot <- downloadHandler(
    filename = function(){
      paste0(input$fig_name, 
             "_shinyfig_",
             day(now()), "-",
             hour(now()), "-",
             minute(now()), ".png")
    },
    content = function(file) {
      ggsave(file,
             rv$plot_selected_plot,
             device = "png",
             width = 10, # width and height are fixed -- cannot seem to get reactive input working
             height = 8,
             units = "cm",
             dpi = 300,
             scale = 1)
    }
  )
})

```

### 2. View model output

__Raw output of a linear model__ using your selected variables. Interpretation is an exercise left to the viewer. 

```{r show-model}

observe({
  if (rv$x_term != "not_sel" & rv$y_term != "not_sel") {
    x_term <- sym(rv$x_term)
    y_term <- sym(rv$y_term)

    # data for model
    raw_dat <- relevant_dat() %>%
      filter(is.na(!!x_term) == FALSE,
             is.na(!!y_term) == FALSE)

    # make model formula
    user_specified_formula <- as.formula(paste0(rv$y_term,
                                                " ~ ",
                                                rv$x_term))
    
    # use model formula in basic lm
    model1 <- lm(user_specified_formula,
                 data = relevant_dat())

    # prepare model table for display
    output$table_model <- renderPrint({
      summary(model1)
    })
    
    # prepare user formula for display (to reference against output table)
    output$text_formula <- renderPrint({
      paste0("User specified formula: ", 
             rv$y_term, 
             " ~ ",
             rv$x_term)
    })
    
  }
}) %>% bindEvent(input$cont_pred_chce,
                 input$cate_pred_chce,
                 input$cont_resp_chce)

# show model output + formula
verbatimTextOutput("text_formula")
verbatimTextOutput("table_model")

```

### 3. View selected data

__Table 1.__ Summarised data for your selected variables. 

_(Remember that summary data will not exist for a scatter plot --- you don't need to summarise the data for that plot type!)_

```{r show-summarised-dat}

(downloadButton("save_summary_dat",
                "Save your summarised data..."))

DTOutput("table_summary_plot_dat")

reactive({
  output$save_summary_dat <- downloadHandler(
    filename = function(){
      paste0(input$fig_name, 
             "_shinysummarydat_",
             day(now()), "-",
             hour(now()), "-",
             minute(now()), ".csv")
    },
    content = function(file) {
      write_csv(rv$summary_plot_dat,
                file)
    }
  )
})


```
<br>
__Table 2.__ Raw data for your selected variables. 

```{r show-raw-dat}

(downloadButton("save_raw_dat",
                "Save your raw data..."))

DTOutput("table_raw_plot_dat")

reactive({
  output$save_raw_dat <- downloadHandler(
    filename = function(){
      paste0(input$fig_name, 
             "_shinyrawdat_",
             day(now()), "-",
             hour(now()), "-",
             minute(now()), ".csv")
    },
    content = function(file) {
      write_csv(rv$raw_plot_dat,
                file)
    }
  )
})

```

### Metadata information

```{r show-metadata}

metadata_for_dt <- metadata %>% 
  rename("Dataset" = dataset,
         "Variable name (for R)" = variable,
         "Variable name (for humans)" = illegal_variable,
         "Units" = units,
         "Description" = description) %>% 
  select(-c(cate_cont_type, pred_resp_type))

output$table_metadata <- renderDT(
  metadata_for_dt,
  rownames = FALSE,
  options = list(pageLength = nrow(metadata),
                 dom = "ft")
)

DTOutput("table_metadata")

```


```{r exit-and-scratch}

# knitr::knit_exit()
# 
# rv$x_term <- "site_type"
# rv$y_term <- "dbh"
# 
# isolate(rv$x_term)
# isolate(rv$y_term)
# 
# formula_test <- as.formula(paste0(isolate(rv$y_term),
#                                   " ~ ",
#                                   isolate(rv$x_term)))
# 
# lm(formula_test,
#    data = relevant_dat)

### x. working on a map

# read country and state shapefiles
# aus_country <- st_read("global/aus_mapping/gadm36_AUS_0.shp")
# aus_states <- st_read("global/aus_mapping/gadm36_AUS_1.shp")
# 
# crop_aus_country <- st_crop(aus_country,
#                             xmin = 113, xmax = 154,
#                             ymin = -11, ymax = -44)
# 
# crop_aus_states <- st_crop(aus_states,
#                            xmin = 153.38, xmax = 153.65,
#                            ymin = -27.37, ymax = -27.51)
# 
# # K'gari crop for full map
# crop_aus_states_upd <- st_crop(aus_states,
#                                xmin = 153.375, xmax = 153.46,
#                                ymin = -27.365, ymax = -27.51)
# 
# fig_colours <- c("map_fill" = "grey80",
#                  "map_site_labels" = "grey70",
#                  "map_points" = "black",
#                  "map_panel_border" = "grey80")
# 
# main_map_upd <- ggplot() +
#   geom_sf(data = crop_aus_states_upd,
#           colour = NA,
#           fill = fig_colours["map_fill"]) +
#   coord_sf(crs = "+proj=longlat +datum=WGS84",
#            label_graticule = "ES",
#            expand = FALSE) +
#   geom_point(data = tree_full,
#              aes(x = lon,
#                  y = lat),
#              size = 1, colour = fig_colours["map_points"]) +
#   geom_text(data = tibble(x = c(153.435,
#                                 153.427,
#                                 153.42),
#                           y = c(-27.417,
#                                 -27.444,
#                                 -27.467),
#                           label = c("Amity\nPoint",
#                                     "Cooroon\nCooroonpah\nCreek",
#                                     "Myora\nSprings")),
#             aes(x = x, y = y, label = label),
#             colour = fig_colours["map_site_labels"],
#             vjust = "middle", hjust = "right",
#             size = unit(3.5, units = "points")) + # originally 3.5
#   geom_text(data = tibble(x = c(153.45,
#                                 153.45),
#                           y = c(-27.41,
#                                 -27.475),
#                           label = c("high\nenergy",
#                                     "low\nenergy")),
#             aes(x = x, y = y, label = label),
#             fontface = "italic",
#             vjust = "middle", hjust = "middle",
#             size = unit(3.5, units = "points")) + 
#   geom_segment(aes(x = 153.45, 
#                    y = -27.415, 
#                    xend = 153.45, 
#                    yend = -27.47),
#                size = 0.5,
#                lineend = "butt", linejoin = "mitre",
#                arrow = arrow(length = unit(0.5, "cm"))) +
#   scale_x_continuous(name = "",
#                      breaks = c(153.41, 153.44)) +
#   scale_y_continuous(name = "") +
#   annotation_scale(location = "br",
#                    width_hint = 0.25,
#                    text_cex = 0.8) +
#   theme_hill_map() %+replace%
#   theme(axis.text.x = element_text(margin = margin(2.5, 0, 0, 0)),
#         text = element_text(colour = fig_colours["text"]),
#         axis.ticks = element_line(colour = fig_colours["axis_ticks"]),
#         panel.background = element_rect(fill = fig_colours["background"],
#                                         colour = NA),
#         plot.background = element_rect(fill = fig_colours["background"],
#                                        colour = NA),
#         panel.border = element_rect(fill = NA,
#                                     colour = fig_colours["map_panel_border"]),
#         axis.line = element_line(colour = fig_colours["axis_line"]))

```
